---
title: "Power Calculation: DoD Grant Proposal"
author: "R. Noah Padgett"
date: "5/28/2020"
output: html_document
---


```{r}

set.seed(20200518)
library(ggplot2)
library(dplyr)
library(kableExtra)

source("code/utility_functions.R")

```

# Design Overview

Grant Title:
**Obesity induced immunoproteosome mediators in the pathogenesis of the acute respiratory distress syndrome**

The design is a full factorial of three factors:

1. Diet (LFS vs. HFS)
2. ARDS Protocol (Sham Condition vs. LPS Administration)
3. Inhibitor (Administered ONX-0914 vs. Saline)

Therefore, there are $2\times 2\times 2 = 8$ experimental groups.

# Run Sample Size Estimation Simulation

We used an effect size of $\eta^2 = 0.313$ for the effect of ARDS protocol from previous observations.
The effect size of diet and the interaction of diet and ARDS protocol were set to half the effect size of the ARDS protocol and varied across replications.
We were unsure of the magnitude of the the other effects so these were varied as well but set to a value of 0.

```{r}

# effect size = 0.313
# convert to d (i.e., beta)
eta = 0.313
d <- r2d(eta)
d

# gamma: significance value
# use Bonferroni correction
gamma = 0.05/7

NREP <- 10000
N <- seq(5, 20, 1)
sim.out <- data.frame(matrix(nrow=length(N)*NREP, ncol=6))
colnames(sim.out) <- c("iter", "N", "Ntotal", "r2", "pvalue", "power")
i <- j <- rep <- 1

for(i in 1:length(N)){
  for(rep in 1:NREP){
    # group IDs
    Nt <- N[i]*8
    n <- N[i]
    A <- c(rep(1,n*4),rep(0,n*4))
    B <- rep(c(rep(1, n*2), rep(0, n*2)),2)
    C <- rep(c(rep(1, n), rep(0, n)),4)
    
    dat <- data.frame(A, B, C)
    
    # Generate regression weights for "off factors"
    b <- c(d, d + rnorm(1, 0, 0.25),
           rnorm(1, 0, 0.25),
           d + rnorm(1, 0, 0.25),
           rnorm(3, 0, 0.25))
    
    # data model
    # y = beta0 + beta1*A + beta2*B + beta3*C + beta4*A:B + beta5*A:C + beta6*B:C + beta7*A:B:C
    e <- rnorm(Nt, 0, 1) # random error
    # simulate data
    dat$y = b[1]*A+b[2]*B+b[3]*C+b[4]*A*B+b[5]*A*C+b[6]*B*C+b[7]*A*B*C + e
    # run ANOVA
    fit <- aov(y~A*B*C, data=dat)
    SS <- summary(fit)[[1]]
    r2 <- SS[1,2]/(sum(SS[2:8,2]))
    
    sim.out[j, 1] <- rep
    sim.out[j, 2] <- N[i]
    sim.out[j, 3] <- Nt
    sim.out[j, 4] <- r2
    sim.out[j, 5] <- SS[1, 5]
    sim.out[j, 6] <- as.numeric(SS[1, 5] <= gamma &
                                  SS[2, 5] <= gamma &
                                  SS[4, 5] <= gamma)
    
    j <- j + 1
  }
}

```

## Summarize Simulation

```{r simulation-summary}

plot.dat <- sim.out %>%
  group_by(N) %>%
  summarise(Power = mean(power),
            eta_squ = mean(r2))

kable(plot.dat, format="html", digits=2) %>%
  kable_styling(full_width = T)

```

## Power Curve of Results

THe Results indicates that we need at least 17 rats per group to reach a power of 0.8 to detect all effects of interest.

```{r power-curve}

p <- ggplot(plot.dat, aes(x=N, y=Power))+
  geom_point()+
  geom_hline(yintercept = 0.8, colour="red")+
  labs(x="Number of Rats per Experimental Group",
       y="Estimated Power",
       title="Power Curve of Experimental Design",
       subtitle="17 Rats per Group Needs for Power >= 0.8")+
  scale_x_continuous(breaks=seq(5,20,1))+
  scale_y_continuous(breaks = seq(0,1, 0.2), limits = c(0,1))+
  theme_classic()
p


```


